
\chapter{Dynamical Obstacle Avoidance Framework}
\label{chapter:framework}
Current collaborative robot arms allow more flexible work cells, where they safely collaborate with human operators augmenting productivity in tasks difficult for traditional automation. However,current solutions for safe interaction simply stopping the robot motion  when a collision is detected. This reduces the productivity in an operational setup in which unintended, safe collisions can happen often. Active contact evasion by the robot arm is desirable so that the production process continues despite regular interferences and path obstructions. In the Factory-in-a-day project dynamic collision avoidance technologies have been developed, including a proximity-sensing robot skin, a reactive path-planning solution and a motion control framework based on proximity sensing which is one of the contributions of this  thesis. These technologies have been integrated into a dynamic-obstacle avoidance framework successfully tested in simulation and laboratory set-ups. This chapter presents the obstacle avoidance solution that is currently being implemented with this framework for a collaborative pick and place application prototype. 

Path planning and motion control are usually formalized as separate problems in robotics though both the problems fundamentally solves what a robot should do next. High dimensional configuration spaces, changing environment and uncertainties does not allow to plan real time motion ahead of time requiring a controller to execute the planned trajectory. The fundamental inability to unify both these problems has led to handle the planned trajectory amidst perturbations and unforeseen obstacles using various trajectory execution and deformation mechanisms. Designing an appropriate architecture to handle the information flow between the control and planning components is also not so trivial. The reactive control using proximity sensors developed is purely an application of the state of the art kinematic Hierarchical Quadratic Programming(HQP) solver to combine both path planning and reactive control within the dynamic collision avoidance framework.

\input{doa/01_intro}
\input{doa/02_framework}
\input{doa/02_soa}
% \input{doa/02_methods}
\input{doa/03_tests}
\input{doa/04_conclusion}
