\ifdefined\included
\else
\documentclass[a4paper,11pt,twoside]{StyleThese}
\include{formatAndDefs}
\sloppy
\begin{document}
\fi


\chapter*{Introduction}
\addstarredchapter{Introduction} 
Autonomy is undoubtedly the main goal in Robotics. We just want the robots to be self-reliant and the backing technologies to be as generic as possible. Every incremental research in robotics that happens directly or indirectly strives to automatize certain processes or mechanisms that solves a problem of physical interaction efficiently and effectively. Though it is controversial to talk about the need of such autonomy in robots, it is an undeniable fact that the research tries to solve robotic problems using algorithms with genericness and self-reliance. Autonomy is not only about the robot having higher awareness about itself and the environment in ideal conditions but the other way around. Developing algorithms with an ability to change, adapt or be flexible is the core problem of establishing autonomy in robots. 

Industrial robots work in static settings and they literally do what it is programmed for specific use-cases, though the robots are flexible enough to be programmed for a variety of tasks. This research work makes an attempt to address these issues that separate both these settings in a profound way with special focus on uncertainties. Practical impossibilities of precise sensing abilities lead to a variety of uncertainties in scenarios where the robot is mobile or the environment is dynamic. This work focuses on developing smart strategies to improve the ability to handle uncertainties robustly in humanoid and industrial robots. The contributions of this thesis is listed as follows:
\begin{itemize}
    \item Develop a strategy to model uncertainties of the inertial parameters of a robot for a humanoid robot specific to balancing task scenarios.
    \item Develop a dynamic obstacle avoidance framework proposed for industrial robots equipped with skin sensors for reactivity. 
\end{itemize}
The motivations behind these contributions come from the fact the real life uncertainties affect the robot performance and it is necessary to develop the capabilities to handle them inherently in the technologies we develop.

\input{industrial_robots.tex}
\section{Humanoids in Real Situations}
DARPA Challenge has shown us how the robots perform in reality in spite of all the technologies working in controlled environment or in simulation. There are various aspects to these  robot failures.
\begin{itemize}
    \item \textit{Robot Design}: The mechanical design of the robot has influence on the kind of
    failures it encounters with the environment. Walking on a flat surface is different from rough surfaces and an appropriate mechanical design is necessary to eliminate failures.

    \item \textit{Behavior Design}: The generated behaviors have to be robust to hardware failures. The kind of behaviors we choose have a big influence on the failures as well. This means robot should be handle variations in the task making the behavior robust. The whole body motion is also physically not depending on the external surface contacts. The robots usually have little or no ability to locomote using external contacts. For instance, it is very natural to hold the railings to climb up the stairs which reduces a significant amount of actuation in the joints.
    \item \textit{Human Robot Interaction}: When the human wants to control the robot in all the levels and switch between the modes easily, then it is not interaction anymore but intervention. Also trying to perform tasks as soon as possible puts the robot in a dynamic disadvantage.       
    \item \textit{Software Engineering}: Eliminating programming errors completely is not yet achieved right now due to undetected bugs resulting in abnormal behaviors. 
    
    \item \textit{Error Detection}: Error detection as early as possible can avoid robot falls. 'Push-Recovery' is one such mechanism to handle external perturbations but handling robot generated errors are given less importance. Just aborting the current behavior after detecting errors works but there is a necessity to handle them systematically.  
    \item \textit{Planning and Control}: Changing dimensions introduces command discontinuities in the control. Interaction with the environment creates a variety of structural changes in the kinematics and dynamics of the robot. Switching behaviors happens due to contact changes, singularities, hitting the joint limits and the QP can't handle this change properly resulting in discontinuities in control. There is also no output feedback to the QP program and ineffective use of foot torque/force sensors for walking. Constraint based inverse kinematics is still challenging to eliminate undesired motions of the rest of the robot links while you execute the intended action in the task space.   
\end{itemize}
The above aspects shows the practical reality of the humanoid robots and the necessity of various components to function appropriately to be robust to failures. Balancing of legged robots is an essential safety constraint and the ability to handle failures is the motivation to focus on robust   balancing under inertial uncertainties of the robot model.  



\section{Factory in a Day project}
Robot Automation has been into existence since the 1960s and has seen a lot of technological advancements but they are still challenged by installation time and cost involved in setting up robots specific to the functional needs in factories. There is a lot of risk involved in such investment making it economically less attractive to smaller companies. Moreover, the factory setups are fixed environments with hard coded settings of millimeter precision so that they are perfectly under control at all times. Floating robot bases introduces errors and uncertainties which requires more robust components to handle factory scenarios. There is also a big concern in safety of human beings working in robot environments. Though recent technologies are focused towards collaborative robot control where human beings and robots can work together, these algorithms are either completely not matured to be used in factories or not yet known in the industrial community. This thesis is funded by Factory-in-a-Day project which focuses on these issues mentioned above.

Factory-in-a-Day project puts forward the idea of minimizing the installation time and the cost involved from several months to just a single day. The project focuses on the following aspects of 
robotics which is synonymous to the steps taken in a day to install a robotic setup. 


\begin{itemize}
\item Standardized procedures to design 3D printed custom parts which are usually attached with an existing robot arm and grippers using novel templates to minimize the time taken.
\item The flexibility to be placed in factories without any alteration where intelligent self-calibration and an adaptive framework that helps to connect with the existing machineries in ease.
\item Use rapid teaching to program production tasks in the setup from a rich set of learnable skills
for applications like mould finishing, welding and assembly.
\item Visually intuitive tools for the workers in the factory to assess robot's behavior motion behaviors. Augmented reality can be used to visualize the robot's intended path by wearing a glass to be aware of its activity to be collaborative in nature.
\item A mature human-robot interaction framework to allow the humans collaborate with robots in an unfenced workspace using a dynamic obstacle avoidance framework with proximity skin sensors and 
reactive path planning and control algorithms.
\end{itemize}
The above aspects along with proposed certification procedures and a complete focus on manufacturing
industry can radically change the automation sector. The work done in this thesis orients in the direction of the last aspect which is human-robot collaboration.


\section{Chapter Organization}

\section{Publications}
\begin{itemize}
\item Giftsun, N., Prete, A. Del, & Lamiraux, F. (2017). \textit{Robustness to 
Inertial Parameter Errors for Legged Robots Balancing on Level Ground.} International Conference on Informatics in Control, Automation and Robotics (ICINCO).
\item Bharatheesha, M., Giftsun, N., Hernandez, C., Bergner, F., Dean-leon, 
E., & Cheng, G. (2017). \textit{Dynamic Obstacle Avoidance for Collaborative 
Robot Applications.} IEEE International Conference on Robotics and 
Automation (ICRA).
\ifdefined\included
\else
\bibliographystyle{acm}
\bibliography{These}
\end{document}
\fi