% Choose the language of your thesis passing 'french' or 'english' as
% \documentclass option.
% Note1: The 'page de garde' will always be written in French.
% Note2: You will have an error if you change the language of the document and
%        compile it without cleaning the auxiliary files. Compiling it again
%        should solve the problem.
\documentclass[english,a4paper,11pt,twoside,table]{StyleThese}
\newcommand{\included}{}

\include{formatAndDefs}
\sloppy
\begin{document}
    
\include{page_de_garde}

\makeflyleaf

% \cleardoublepage

\dominitoc

\pagenumbering{roman}

% \cleardoublepage

% Here you can see an example of how to create text conditioned by the language
% variable. The \iftoggle command:
%
%   \iftoggle{ThesisInEnglish}{%
%   <your-text-in-english>
%   }{%
%   <your-text-in-french>
%   }
%
% will compile only one of the two blocks, depending on the variable you set at
% the beginning of this document. Language selection is managed this way in the
% formatAndDefs.tex file. You too can create sections of your thesis that is
% language dependend this way, although you probably won't need it. Another use
% of \iftoggle can be found at the end of this file.
\iftoggle{ThesisInEnglish}{%
\section*{Acknowledgments}
}{%
\section*{Remerciements}
}
A faire en dernier :-) 

\begin{vcenterpage}
\iftoggle{ThesisInEnglish}{%
\section*{Abstract}
}{%
\section*{Resume}
}
 Amidst a lot of research in motion planning and control in concern with robotic applications, the mankind has never reached a point yet, where the robots are perfectly functional and autonomous in dynamic settings. Though it is controversial to discuss about the necessity of such robots, it is very important to address the issues that stop us from achieving such a level of autonomy. Industrial robots have evolved to be very reliable and highly productive with more than 1.5 million operation robots in a variety of industries. These robots work in static settings and they literally do what it is programmed for specific usecases, though the robots are flexible enough to be programmed for a variety of tasks. This research work makes an attempt to address these issues that separate both these settings in a profound way with special focus on uncertainties. Practical impossibilities of precise sensing abilities lead to a variety of uncertainties in scenarios where the robot is mobile or the environment is dynamic. This work focuses on developing smart strategies to improve the ability to handle uncertainties robustly in humanoid and industrial robots. First, we focus on a dynamical obstacle avoidance framework proposed for industrial robots equipped with skin sensors for reactivity. Path planning and motion control are usually formalized as separate problems in robotics, though both fundamentally address the same issue. High dimensional configuration spaces, changing environment and uncertainties do not allow to plan real-time motion ahead of time requiring a controller to execute the planned trajectory. The fundamental inability to unify both these problems has led to handle the planned trajectory amidst perturbations and unforeseen obstacles using various trajectory execution and deformation mechanisms. The proposed framework uses 'Stack of Tasks', a hierarchical controller using proximity information with a point cloud based reactive path planner to avoid obstacles. Experiments are performed in UR5 robot to check the validity of the framework both in simulation and real robot. 

 Second, we focus on a strategy to model uncertainties of the inertial parameters of a robot for a humanoid robot specific to balancing task scenarios. Model-based control has become more and more popular in the legged robots community in the last ten years. The key idea is to exploit a model of the system to compute precise motor commands that result in the desired motion. This allows to improve the quality of the motion tracking, while using lower feedback gains, leading so to higher compliance. However, the main flaw of this approach is typically its lack of robustness to modeling errors. In this paper we focus on the robustness of inverse-dynamics control to errors in the inertial parameters of the robot. We assume these parameters to be known, but only with a certain accuracy. We then propose a computationally-efficient optimization-based controller that ensures the balance of the robot despite these uncertainties. We used the proposed controller in simulation to perform different reaching tasks with the HRP-2 humanoid robot, in the presence of various modeling errors. Comparisons against a standard inverse-dynamics controller through hundreds of simulations show the superiority of the proposed controller in ensuring the robot balance.


\end{vcenterpage}

\tableofcontents

\printnomenclature
% Use \mtcfixnomenclature below if you have a glossary (added with
% \printnomenclature above) and you're see a shift in the mini-table of
% contents at the begining of each chapter (example: no mini-toc in chapter 1;
% mini-toc of chapter 1 appearing in chapter 2; and so on).
%
% You should not use \mtcfixnomenclature if you have no glossary (that means,
% if you don't use \printnomenclature or if your glossary is empty).
\mtcfixnomenclature

\mainmatter

\include{Introduction}
%\include{Chapter1}
%\include{Chapter2}
\include{doa/doa}
\input{rc_inertial_params/RobustController}
\include{Conclusion}
% \appendix
% \include{Annexe1}

\bibliographystyle{StyleThese}
%\bibliographystyle{plain}
\bibliography{These}

% \cleardoublepage
% \begin{vcenterpage}
% \noindent\rule[2pt]{\textwidth}{0.5pt}
% \\
% \iftoggle{ThesisInEnglish}{%
% {\large\textbf{Abstract:}}
% }{%
% {\large\textbf{Résumé :}}
% }
% resume
% \iftoggle{ThesisInEnglish}{%
% {\large\textbf{Keywords:}}
% }{%
% {\large\textbf{Mots clés :}}
% }
% mots, clefs
% \\
% \noindent\rule[2pt]{\textwidth}{0.5pt}
% \end{vcenterpage}

\end{document}
