\ifdefined\included
\else
\documentclass[a4paper,11pt,twoside]{StyleThese}
\include{formatAndDefs}
\sloppy
\begin{document}
\fi
\chapter*{Introduction}
\addstarredchapter{Introduction} 
Autonomy is undoubtedly the main goal in Robotics. We would like to develop robots which are self-reliant with supporting technologies to be as generic as possible. Every incremental research in robotics  directly or indirectly strives to automatize certain processes or mechanisms that solve a problem of physical interaction efficiently and effectively. Though it is controversial to talk about the need of such autonomy in robots, it is an undeniable fact that the research tries to solve robotic problems using algorithms with genericness and self-reliance. Autonomy is not only about robot having higher awareness and control on itself in ideal or pre-designed environments but also the ability to cope up with new and dynamic environments with ease. Developing algorithms with an ability to change, adapt or be flexible is the core problem of establishing autonomy in robots. 

In this thesis, we address two specific problems which comes under this futuristic vision of establishing advanced autonomy in robot control:\textit{Variability} in environment and \textit{Uncertainty} in robot model. By the term \textit{Variability}, we refer to environments of being in state of inconsistency or lack of fixed pattern for which an appropriate algorithm should be flexible enough to adapt without compromising on the final goal. Whereas by the term \textit{Uncertainty}, we refer to the incomplete knowledge of qualitative or quantitative aspects of a robot requiring a control algorithm to succeed in goals even without complete knowledge. Though these two terms look like they are very related to each other and it might even mean the same in some scenarios, they are fundamentally different and need to be handled differently. Both these problems are addressed in two different robotic platforms motivated by the current need of the hour in robotics research and development.

The first part of the thesis presents the development of a controller framework in industrial robots to handle variability in environment. Industrial robots work in static settings and they literally do what it is programmed for specific use-cases, though the robots are flexible enough to be programmed for a variety of tasks. This work makes an attempt to address these issues that separate both these settings in a profound way with special focus on dynamic obstacle avoidance where the robots encounter unforeseen obstacles while executing a desired task. 
The second part of the thesis presents the development of a robust controller for humanoid robot to ensure balance even when there is no absolute knowledge of the robot model. Lack of precise sensing abilities lead to a variety of uncertainties in robot model and the environment. In legged robots, due to the tree-like kinematic chain and the dynamic complexity, it is necessary to approximate the robot to a simple model for balance control. Unfortunately, it is quite sensitive to uncertainties of the model for which a robust control strategy is proposed to improve balance control amidst inertial uncertainties of the robot.

This thesis in short focuses on developing smart strategies to improve the ability to handle variability and uncertainties robustly both in humanoid and industrial robots respectively. The contributions of this thesis are:
\begin{itemize}
    \item A strategy to model uncertainties of the inertial parameters of a robot for a humanoid robot specific to balancing task scenarios.
    \item A dynamic obstacle avoidance framework proposed for industrial robots equipped with skin sensors for reactivity. 
\end{itemize}

In this chapter, we introduce the field of both industrial \& humanoid robots and the issues we face to motivate the need to solve variability and uncertainties specific to the context of this thesis. 
\input{chapters/intro/industrial-robots.tex}
\input{chapters/intro/humanoid-robots.tex}
\input{chapters/intro/control-methodologies.tex}

\section{Chapter Organization}
The thesis has two main chapters \ref{chapter:framework} and \ref{chapter:rc} describing the two main contributions of this thesis respectively. Chapter \ref{chapter:framework} presents the dynamic obstacle framework developed on industrial robots for collision avoidance using proximity information from skin sensors and point cloud information from 3d cameras. The Chapter \ref{chapter:rc} presents the strategy used to model inertial uncertainties to perform robust balancing. Chapter 5 concludes the thesis with important comments on the contributions and the possible future work connected to it.
\section{Publications}
\begin{itemize}
\item Giftsun, N., Prete, A. Del, \& Lamiraux, F. (2017). \textit{Robustness to Inertial Parameter Errors for Legged Robots Balancing on Level Ground.} International Conference on Informatics in Control, Automation and Robotics (ICINCO) \textbf{[CONFERENCE]}.
\item Bharatheesha, M., Giftsun, N., Hernandez, C., Bergner, F., Dean-leon, E., \& Cheng, G. (2017). \textit{Dynamic Obstacle Avoidance for Collaborative Robot Applications.} IEEE International Conference on Robotics and Automation (ICRA)\textbf{[WORKSHOP]}.
\end{itemize}
\ifdefined\included
\else
\bibliographystyle{acm}
\bibliography{These}
\end{document}
\fi